Learning to Synthesize Rules for Rule-based Intrusion Detectors

### 2020-03-06 paper/paper.tex
pass on RQ3

\subsubsection{Popularity of http content modifiers.} As previously
mentioned, \percRulesWithContent\ of \suri\ rules involve content
options. Furthermore, nearly half of the rules that use content
options use content modifiers to qualify the search. Http modifiers
typically serve to specify parts of the message payloads that the
input string should be matched against. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} indicates that the string ``keep-alive'' should be
present in the header of the http message\footnote{A keep-alive
  connection attribute indicates that a single tcp connection should
  remain open for multiple http requests/responses.}. Note that, in
this case, the content modifier refers to a field of an http message,
which appears in the payload of a tcp message, which is the target of
the option \CodeIn{content}. Figure~\ref{fig:http-header-example}
illustrates possible content options used to match against certain
strings that appear in an http message. A check mark (respectively,
cross) indicates a match (respectively, no match).

@@

This section evaluates the importance of each proposed heuristic
function. To that end, we evaluated the use of \tname{} without each
of these functions with the goal of checking how much each of them
contributes to the overall performance of the technique. Recall that
the score a function $r$ is given by the weigthed average of the
values given by each heuristic function, \ie{} the score of $r$ is
obtained with the aggregate function $f(r)=\sum_{i}^{N}
w_i*f_i(r)/N;$, where $w_i$ is the weight of heuristic $i$, $f_i(r)$
is the score of heuristic $i$ on $r$, and $N$ is the number of
heuristic---four, in our case. To assess the value of each heuristic
function, we assigned its weight to zero---effectively, disabling the
function---and compared the result with \tname\ using default weights,
obtained with the procedure described on
Section~\ref{sec:weight-optimization}.

Table~\ref{table:weights} shows results. Column ``Name'' shows the
name of the attack and column ``All'' shows the result of \tname\ with
all heuristic functions enabled; as in column ``Rank (step 3)'' from
Table~\ref{table:results}. The remaining columns show results obtained
by running the same experiment as in Section~\ref{sec:answer-rqone},
but supressing one heuristic function. An up arrow indicates that,
considering that particular data point, the performance of \tname{} on
that configuration is worse when compared to \tname's default
configuration. A down arrow, appearing highlighted in the table,
indicates the contrary.

\Mar{--------------------------------- estou aqui}

\setlength{\tabcolsep}{5pt}
\begin{table}[h!]
  \small
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    All &
    -f2 &
    -f3 &    
    -f4 &
    -f5 \\ 
    Black Nurse & 2 & 2 & 2 & 2 & 7$\uparrow$ \\    
    Ping Scan & 2 & 2 & 2 & 2 &  8$\uparrow$\\
    SYN Flood & 2 & 2 & 2 & 2 & 2 \\
    UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
    %%UDP Flood & - & - & - & - & - \\

    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 3 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ & 3 \\
    iisadmin access & 1 & 1 & 1 & 8$\uparrow$ & 1 \\        
    JBoss JMX & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ & 45 \\
    Script tag in URI & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ & 1 \\
    WeBid LFI & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ & 10 \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 2 & 2 & 2 & 17$\uparrow$ & 2 \\
    .idq access & 2 & 2 & 2 & 12$\uparrow$ & 2 \\
    /system32/ in URI & 4 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$ & 4 \\
    Basebuilder LFI & 34 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$ & 34 \\
    Tomcat Snoop Access & 32 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$ & 32 \\
    \bottomrule
  \end{tabular}
\end{table}

Note that the number of up arrows is relatively high and the magnitude
of the degradation can be very high in some cases. Basebuilder and
Tomcat are two cases where results get much worse when supressing one
of the functions. Also note that the number of down arrows is
relatively small. Recall that the optimization of weights looks for
global optimum and cannot avoid those cases. Finally, note that every
function contributed. Some functions contributed in more cases than
others, namely \Fix{xxx}, and one of the functions, namely \Fix{xxx},
only contributed to the performance of \tname{} on attacks involving
non-http protocols.

\begin{center}
\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
..........................
\end{tcolorbox}
\end{center}

@@

This section evaluates sensitivity of \tname{} to the number of
attacks provided on input. Recall that \tname\ uses one example attack
to generate the seed rule (at the Reverse Engineering step) and the
remaining attacks to evaluate soundness of the rule (at the Rules
Creations step). It is expected that the number and variety of
attacks---of the same kind---plays an important role in the
performance of the technique. To answer this research question, we
selected the cases with the highest number of examples of attacks (see
coolumn ``\#~Vars'' on Table~\ref{table:results}), varied the number of
examples, and measured the position of the golden rule at the
rank. For each number of variants $N$, we sampled five sets of size
$N$ to provide as input to the technique.

Table~\ref{fig:impact-number-attacks} shows
results. \Fix{...elaborate...}

\begin{figure}[t!]
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.4,trim=20 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.4,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \caption{scripttag}    
  \end{subfigure}%
\caption{Impact of number of positive examples.}
\label{fig:impact-number-attacks}
\end{figure}

>>>>

\subsubsection{\label{sec:popularity-content-modifiers}Popularity of http content modifiers.} As previously
mentioned, \percRulesWithContent\ of \suri\ rules involve content
options. Furthermore, nearly half of the rules that use content
options use content modifiers to qualify the search. Http modifiers
typically serve to specify parts of the message payloads that the
input string should be matched against. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} indicates that the string ``keep-alive'' should be
present in the header of the http message\footnote{A keep-alive
  connection attribute indicates that a single tcp connection should
  remain open for multiple http requests/responses.}. Note that, in
this case, the content modifier refers to a field of an http message,
which appears in the payload of a tcp message, which is the target of
the option \CodeIn{content}. Figure~\ref{fig:http-header-example}
illustrates possible content options used to match against certain
strings that appear in an http message. A check mark (respectively,
cross) indicates a match (respectively, no match).

@@

RQ2 evaluates the importance of each proposed heuristic function. To
answer that question, we evaluated the use of \tname{} without each of
these functions with the goal of checking how much each of them
contributes to the overall performance of the technique. Recall that
the score a rule $r$ is given by the weigthed average of the values
given by each heuristic function, \ie{} the score of $r$ is obtained
with the function $f(r)=\sum_{i}^{N} w_i*f_i(r)/N$, where $w_i$ is the
weight of heuristic $i$, $f_i(r)$ is the score of heuristic $i$ on
$r$, and $N$ is the number of heuristics considered---four, in our
case. To assess the value of each heuristic function, we assigned its
weight to zero---effectively, disabling the function---and compared
the result with \tname\ using default weights, obtained with the
optimization procedure described on
Section~\ref{sec:weight-optimization}.

Table~\ref{table:weights} shows results. Column ``Name'' shows the
name of the attack and column ``base'' shows the result of
\tname\ with all heuristic functions enabled, as in column ``Rank
(step 3)'' from Table~\ref{table:results}. The remaining columns show
results obtained by running the same experiment, as on
Section~\ref{sec:answer-rqone}, but supressing one heuristic
function. The number corresponds to the order the heuristic function
is introduced on Section~\ref{sec:ranking}. The symbol of an up arrow
($\uparrow$) indicates that the performance of \tname{} on that
configuration is worse when compared to \tname's default
configuration. A down arrow ($\downarrow$), appearing highlighted in
the table, indicates the contrary.

\setlength{\tabcolsep}{5pt}
\begin{table}[h!]
  \small
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 2 & 2 & 2 & 2 & 2 \\
    UDP Teardrop & 4 & 4 & 4 & 4 & 4 \\
    %%UDP Flood & - & - & - & - & - \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 3 & 3 & 2\cellcolor{lightgray}$\downarrow$ & 12$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 8$\uparrow$\\        
    JBoss JMX & 45 & 45 & 5\cellcolor{lightgray}$\downarrow$ & 45 & 111$\uparrow$ \\
    Script tag in URI & 1 & 1 & 4$\uparrow$ & 1 & 8$\uparrow$ \\
    WeBid LFI & 10 & 10 & 77$\uparrow$ & 28$\uparrow$ & 32$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 2  & 2 & 2 & 2 & 17$\uparrow$\\
    .idq access & 2  & 2 & 2 & 2 & 12$\uparrow$ \\
    /system32/ in URI & 4  & 4 & 20$\uparrow$ & 1\cellcolor{lightgray}$\downarrow$ & 58$\uparrow$\\
    Basebuilder LFI & 34  & 34 & 117$\uparrow$ & 57$\uparrow$ & 183$\uparrow$\\
    Tomcat Snoop Access & 32  & 32 & 828$\uparrow$ & 2\cellcolor{lightgray}$\downarrow$ & 2564$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

Note that the number of up arrows is relatively high and the magnitude
of the degradation can be very high in some cases. Basebuilder and
Tomcat are two cases where results get much worse when supressing one
of the functions. Also note that the number of down arrows is
relatively small. Recall that the optimization of weights looks for
global optimum and cannot avoid those cases. Finally, note that every
function contributed. Function ``Popularity of content modifiers''
(Section~\ref{sec:popularity-content-modifiers}) was the function that
contributed in more cases whereas function ``Popularity of options''
(see Section~{sec:popularity-options}) only contributed to the
performance of attacks involving non-http protocols.

\begin{center}
\begin{tcolorbox}[enhanced,width=3.3in,center upper,drop shadow southwest,sharp corners]
Every heuristic function contributed to the overall performance of
\tname. The function that builds on the popularity of content
modifiers was the one that affected positively the higher number of
cases we analyzed.
\end{tcolorbox}
\end{center}

@@

This section evaluates sensitivity of \tname{} to the number of
attacks provided on input. Recall that \tname\ uses one example attack
to generate the seed rule (at the Reverse Engineering step) and the
remaining attacks to evaluate soundness of the rule (at the Rules
Creations step). It is expected that the number and variety of
attacks---of the same kind---plays an important role in the
performance of the technique.

To answer this research question, we selected two cases with the
highest number of examples (see column ``\#~Vars'' on
Table~\ref{table:results})---idqaccess and scripttag---and the case
where \tname\ performs worst---jboss. We measured the position of the
golden rule at the rank for an increasing number of of examples
provided to \tname. For each number of variants, we sampled five sets
of that size to provide as input to the
technique. Figure~\ref{fig:impact-number-attacks} shows the results
for each attack we selected. The y-axis of each plot denotes the rank
of the golden rule and x-axis denotes the number of attack examples
used to synthesize rules. Each point in the x-axis is associated with
a boxplot for the distribution of results for the corresponding number
of variants. Note that 1) the number of examples is important, 2) the
choice of examples is important, and 3) a small number of examples
suffices to reach the best results.

\begin{figure}[h!]
  \centering
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.4,trim=20 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.2\textwidth}
    \centering
    \includegraphics[scale=0.4,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \caption{scripttag}    
  \end{subfigure}%
\caption{Impact of number of positive examples.}
\label{fig:impact-number-attacks}
\end{figure}
